---
title: "STAT 331 Final Project: Mystery Concert"
date: now
author: "Mia Sahagian, Suhani Kashyap, Hanna Karam"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
editor: source
execute: 
  echo: true
  message: false
  warning: false
---

## Packages

```{r}
library(tidyverse)
library(knitr)
library(janitor)
library(gt)
library(broom)
library(knitr)
```

# Mystery Concert

**Background**: A mystery music artist is organizing a concert at Cal Poly. The biggest celebrities in the world are dropping hints on social media about who is behind it. As a journalist at Mustang News, you need to follow the hints across Spotify, Apple Music, and global information to discover the mystery artist and let the students at Cal Poly know who’s performing.

## Data Dictionary & Import

The three CSV files used in this lab were downloaded directly from Kaggle and are linked below. Download the CSV files and save them in your directory.

#### Global Music:

This dataset includes 100,000+ music artists from around the world, recording names, genres (pop, rock, jazz, or classical), profile image (in a link), and location (where the artist is from).

Find more info on the data here: <https://www.kaggle.com/datasets/harshdprajapati/worldwide-music-artists-dataset-with-image>

#### Apple Music:

The dataset includes information on various artists/tracks such as artist name, track name, genre, track duration, and release date, among others.

Find more info on the data here: <https://www.kaggle.com/datasets/kanchana1990/apple-music-dataset-10000-tracks-uncovered>

#### Spotify Music:

The dataset includes information about songs such as artist name, album name and release date, as well as audio features such as tempo, danceability, acoustiness, energy, and more.

Find more info on the data here: <https://www.kaggle.com/datasets/solomonameh/spotify-music-dataset?resource=download>

```{r}

global_music <- read.csv("./Global Music Artists.csv")

apple_music <- read.csv("./apple_music_dataset.csv")

spotify_music <- read.csv("./high_popularity_spotify_data.csv")

#if you named the csv differently, edit the file path
```

## Data Cleaning

Before you can discover the mystery artist playing at Cal Poly, you need to clean up the data that you're looking at. The students at Cal Poly are counting on you!

### 1. Clean the Apple Music Data

Make a dataset called `apple_clean` and keep only the variables needed for the investigation: `artistName`, `trackName`, `primaryGenreName`, `releaseDate`, and `trackTimeMillis` . There should be no columns related to censorship, explicitness, and price. Rename these variables if necessary. Edit the `releaseDate` column so it has the date in year-month-day, hours-min-second format; extract the release year from the date, and make the track time be in minutes.

```{r}
apple_clean <- apple_music |>
 select(artist_name = artistName,
        track_name = trackName,
        genre = primaryGenreName,
        release_date = releaseDate,
        track_time_millis = trackTimeMillis) |>
 mutate(release_date = ymd_hms(release_date),
        release_year = year(release_date),
        track_duration_min = track_time_millis / 60000)
```

### 2. Clean the Spotify Music Data

Make the dataset `spotify_clean` and keep only variables related to artist identity, popularity, release date, and audio features. Keep only variables `track_name`, `track_artist`, `track_popularity`, `track_album_name`, `track_album_release_date`, `playlist_genre`, `danceability`, `energy`, `tempo`, and `acousticness`. Create a new column called `release_year` that has the year the track was released. Don't include `track_album_release_date`, just use it to create `release_year`.

::: {.callout-note style="color: green"}
## Hint!

Look into parse_date_time() !
:::

```{r}
spotify_clean <- spotify_music |>
 select(track_name,
        track_artist,
        track_popularity,
        track_album_name,
        track_album_release_date,
        playlist_genre,
        danceability,
        energy,
        tempo,
        acousticness) |>
 mutate(track_album_release_date = parse_date_time(track_album_release_date, 
                                                   orders = c("ymd", "ym", "y")),
        release_year = year(track_album_release_date)) |>
 select(-track_album_release_date)
```

### 3. Clean the Global Music data

Make the dataset `global_clean` and keep only `artist_name`, `country`, and `genre`.

```{r}
global_clean <- global_music |>
 select(artist_name,
        country,
        artist_genre)
```

#### Check in: Check sizes

After cleaning:

-   `apple_clean`: each row represents one Apple Music track, with artist, track, genre, release date, release year, time, and duration.
-   `spotify_clean`: each row represents one Spotify track, with artist, popularity, album, release year, genre, and audio features.
-   `global_clean`: each row represents one global music artist, including artist name, country, and genre.

Apple Music should be 10000 rows by 7 columns, Spotify Music should be 1686 rows by 10 columns, and Global Music should be 101375 rows by 3 columns.

```{r}
kable(tibble(dataset = c("Apple Music", "Spotify Music", "Global Music"),
             rows = c(nrow(apple_clean), nrow(spotify_clean), 
                      nrow(global_clean)),
             columns = c(ncol(apple_clean), ncol(spotify_clean), 
                         ncol(global_clean))),
      caption = 'Verified clean dataset dimensions')
```

## Follow the clues!

### Clue 1: Most common music genre plot

Congrats on finishing the first step and cleaning up the data! You feel a bit hopeless because you don't have any clues, but then, your friend alerts you that Doja Cat just posted a TikTok about the Cal Poly concert, saying that "The mystery artist is definitely from the most common genre".

Your job now is to find the most common genres in the Apple Music data by creating a plot and sorting the top 10 genres from least common to most common. Which genre is most common?

```{r}
apple_clean |>
 count(genre, sort = TRUE) |>
 slice_max(n, n = 10) |>
 ggplot(aes(x = fct_reorder(genre, n), y = n)) +
 geom_col(fill = 'skyblue') +
 labs(title = "Most Common Apple Music Genres",
      x = "Genre",
      y = "",
      subtitle = 'Number of Tracks') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25))
```

The most common music genre in the Apple music data is Pop.

### Clue 2: Promo song length

Hurrah! You're one step closer to finding the mystery artist. There are some talks of there being a certain song that is being used as a promo track, and as a data-driven Mustang News journalist, you know that narrowing down that promo track will help you narrow down the mystery headliner. Justin Bieber accidentally leaks on his Instagram Live that “The song is around four minutes long".

Write a function to classify song length that takes a song duration in minutes as input and outputs if a song is too short, a possible match, or too long. Possible matches will be between 3.5 and 4.5 minutes. Make a new dataset containing all possibilities from apple music (remember to use your genre from the previous clue).

```{r}
song_length <- function(minutes) {
  case_when(minutes < 3.5 ~ "Too short",
            minutes <= 4.5 ~ "Possible match",
            minutes > 4.5 ~ "Too long")}

apple_artists <- apple_clean |>
  mutate(duration_match = song_length(track_duration_min)) |>
  filter(genre == "Pop",
         duration_match == "Possible match")

kable(head(apple_artists),
      caption = 'Apple Music tracks matching Pop genre and song length clues')
```

### Clue 3: Promo song filter

Bad Bunny tweets, "The headlining song at the Cal Poly concert is stuck in my head. It's got a nice groove; it's not too energetic, but it's still dance-worthy."

The hit song played by the mystery artist has medium energy (45%–65%) and medium-high danceability (60%–75%).

Write a data frame function called between_filter() that:

-   takes a data frame, a variable name, a minimum value, and a maximum value,

-   and returns only the rows where the chosen variable falls between the minimum and maximum values (inclusive).

Use your function twice to create a new data frame called `spotify_artists` containing songs with: `energy` between 0.45 and 0.65, and `danceability` between 0.60 and 0.75. Display the first rows of `spotify_artists` using kable().

```{r}
between_filter <- function(df, var, min, max) {
  
  df |>
    filter({{ var }} >= min,
           {{ var }} <= max)
  
}

spotify_artists <- spotify_clean |>
  between_filter(energy, 0.45, 0.65) |>
  between_filter(danceability, 0.60, 0.75)

kable(head(spotify_artists),
      caption = 'Spotify tracks matching the energy and danceability clues')
```

### Clue 4: Combine Datasets

Since such huge names are talking about this mystery artist, you know that the artist has songs on both Apple Music and Spotify. Remove duplicate songs from both `apple_artists` and `spotify_artists`, then create a combined dataset called `final_artists` containing only songs that appear in both datasets.

::: {.callout-note style="color: green"}
## Hint!

Try using distinct() to remove any duplicates of songs!
:::

```{r}
spotify_artists <- spotify_artists |>
  distinct(track_name, track_artist, .keep_all = TRUE)

apple_artists <- apple_artists |>
  distinct(track_name, artist_name, .keep_all = TRUE)

final_artists <- spotify_artists |>
  inner_join(apple_artists,
             by = join_by(track_name == track_name,
                          track_artist == artist_name,
                          release_year == release_year))
```

### Clue 5: Predict Popularity

TMZ leaks that the promo song is a massive hit. However, insiders claim that the song is more popular than you would expect based on its audio characteristics alone. Fit a linear regression model called `popularity_model` that predicts `track_popularity` using: `danceability`, `energy`, `tempo`, `acousticness`. Use `augment()` to add the model's predicted popularity values to `final_artists`. Create a new variable called `popularity_difference` that measures how much more popular a song is than expected. Keep only the three songs with the largest positive popularity difference. Display the resulting data frame using `kable()`.

```{r}
popularity_model <- lm(track_popularity ~ danceability + 
                         energy + tempo + acousticness,
                       data = final_artists)

final_artists <- popularity_model |>
  augment(newdata = final_artists) |>
  mutate(popularity_difference = track_popularity - .fitted) |>
  slice_max(popularity_difference, n = 3)

final_artists |>
  head() |>
  kable(caption = 'Top 3 songs with the largest positive popularity difference')
```

### Clue 6: International Time Travel

A fan notices that all of these hints were posted at some strange times in Pacific Time. You suspect that the mystery artist is actually NOT from the United States. You notice that most of these hints are posted when the UK is up and at 'em.

Then, Drake tweets "Yoo, this Cal Poly headliner got me mad nostalgic, reminds me of my 'Kiki do you love me' days. The promo song is one of the best songs from the 2010s.

Join the remaining potential artists to the global artist data and filter for artists from the United Kingdom. Further narrow the results to songs released during the 2010s. Complete the final filter based on these clues, and display your results using the `gt()` function.

```{r}
final_artists <- final_artists |>
  left_join(global_clean,
            by = join_by(track_artist == artist_name)) |>
  filter(str_detect(country, "United Kingdom"),
         release_year >= 2010,
         release_year < 2020)

final_artists |>
  select(track_name,
         track_artist) |>
  gt() |>
  tab_header(title = "Final Mystery Concert Artist and Song Reveal") |>
  cols_label(track_name = "Song",
             track_artist = "Artist")
```

::: callout-note
## Result

After following all six clues and combining evidence from Apple Music, Spotify, and Global Music data, the artist and their headlining song is ... One Direction headlining with Night Changes!!

You can now write your front page Mustang News article!
:::
