---
title: "STAT-331-Final-Project"
date: now
author: "Mia Sahagian, Suhani Kashyap, Hanna Karam"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
editor: source
execute: 
  echo: true
  message: false
  warning: false
---

## Style to dos

```{r}
# Rewrite all questions + make it more fun like this artists said this type pls
# Make sure all questions include a clear description on what to do for that question
# I lowk labeled it suspects from instinct but maybe a different word would be better so maybe rename data
#but if you rename the data make sure it gets changed everywhere
# Add any callouts/that purple thing that looks like code out of code
# Check with her labs to see if anything else should be included for style
#pls check my code and logic
#I just did displays with head() for logic but we can change this idk, kable instead or nothing idk
#when I say high danceability or whatever we need to provide a range instead/also in the question
```

## Packages

```{r}
library(tidyverse)
library(knitr)
library(janitor)
library(gt)
```

# Mystery Concert

**Background**: A mystery music artist is organizing a concert at Cal Poly. The biggest celebrities in the world are dropping hints on social media about who is behind it. As a journalist at Mustang News, you need to follow the hints across Spotify, Apple Music, and global information to discover the mystery artist and let the students at Cal Poly know who’s performing.

## Data Dictionary & Import

#### Global Music:

This dataset includes 100,000+ music artists from around the world, recording names, genres (pop, rock, jazz, or classical), profile image (in a link), and location (where the artist is from).

Find more info on the data here: <https://www.kaggle.com/datasets/harshdprajapati/worldwide-music-artists-dataset-with-image>

#### Apple Music:

The dataset includes information on various artists/tracks such as artist name, track name, genre, track duration, and release date, among others.

Find more info on the data here: <https://www.kaggle.com/datasets/kanchana1990/apple-music-dataset-10000-tracks-uncovered>

#### Spotify Music:

The dataset includes information about songs such as artist name, album name and release date, as well as audio features such as tempo, danceability, acoustiness, energy, and more.

Find more info on the data here: <https://www.kaggle.com/datasets/solomonameh/spotify-music-dataset?resource=download>

```{r}

global_music <- read.csv("./Global Music Artists.csv")

apple_music <- read.csv("./apple_music_dataset.csv")

spotify_music <- read.csv("./high_popularity_spotify_data.csv")
```

## Data Cleaning

Before you can discover the mystery artist playing at Cal Poly, you need to clean up the data that you're looking at. Be quick, the students at Cal Poly are counting on you!

### 1. Clean the Apple Music Data

Make a dataset called `apple_clean` and keep only the variables needed for the investigation: `artistName`, `trackName`, `primaryGenreName`, `releaseDate`, and `trackTimeMillis` . There should be no columns related to censorship, explicitness, and price. Rename these variables if neccessary. Edit the `releaseDate` column sot it has the date in year month day hours min second format; extract the release year from the date, and make the track time be in minutes.

```{r}
apple_clean <- apple_music |>
 select(artist_name = artistName,
        track_name = trackName,
        genre = primaryGenreName,
        release_date = releaseDate,
        track_time_millis = trackTimeMillis) |>
 mutate(release_date = ymd_hms(release_date),
        release_year = year(release_date),
        track_duration_min = track_time_millis / 60000)
```

### 2. Clean the Spotify Music Data

Make the dataset `spotify_clean` and keep only variables related to artist identity, popularity, release date, and audio features. Keep only variables `track_name`, `track_artist`, `track_popularity`, `track_album_name`, `track_album_release_date`, `playlist_genre`, `danceability`, `energy`, `tempo`, and `acousticness`. Create a new column called `release_year` that has the year the track was released. Don't include `track_album_release_date`, just use it to create `release_year`.

::: {.callout-note style="color: green"}
## Hint!

Look into parse_date_time() !
:::

```{r}
spotify_clean <- spotify_music |>
 select(track_name,
        track_artist,
        track_popularity,
        track_album_name,
        track_album_release_date,
        playlist_genre,
        danceability,
        energy,
        tempo,
        acousticness) |>
 mutate(track_album_release_date = parse_date_time(track_album_release_date, 
                                                   orders = c("ymd", "ym", "y")),
        release_year = year(track_album_release_date)) |>
 select(-track_album_release_date)
```

### 3. Clean the Global Music data

Make the dataset `global_clean` and keep only artist name, country, and genre.

```{r}
global_clean <- global_music |>
 select(artist_name,
        country,
        artist_genre)
```

#### Check in: Check sizes

Display with... Apple should be 10000 by 7, spotify 1686 by 10, and global 101375 by 3.

```{r}
kable(tibble(dataset = c("Apple Music", "Spotify Music", "Global Music"),
             rows = c(nrow(apple_clean), nrow(spotify_clean), nrow(global_clean)),
             columns = c(ncol(apple_clean), ncol(spotify_clean), ncol(global_clean))))

```

## Follow the clues!

### Clue 1: Most common music genre plot

Congrats on finishing the first step and cleaning up the data! You feel a bit hopeless because you don't have any clues, but then, your friend alerts you that Doja Cat just posted a TikTok about the Cal Poly concert, saying that "The mystery artist is definitely from the most common genre".

Your job now is to find the most common genres in the Apple Music data by creating a plot and sorting the top 10 genres from least common to most common. Which genre is most common?

```{r}
apple_clean |>
 count(genre, sort = TRUE) |>
 slice_max(n, n = 10) |>
 ggplot(aes(x = fct_reorder(genre, n), y = n, fill = fct_reorder(genre, n))) +
 geom_col() +
 labs(title = "Most Common Apple Music Genres",
      x = "Genre",
      y = "Number of Tracks",
      fill = 'Genre') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25))

#labels from both x-axis and legend...? include or no? Should this be a gradient instead?
```

The most common music genre in the Apple music data is Pop.

### Clue 2: Promo song length

Hurrah! You're one step closer to finding the mystery artist. There are some talks of there being a certain song that is being used as a promo track, and as a data-driven Mustang News journalist, you know that narrowing down that promo track will help you narrow down the mystery headliner. Justin Bieber accidentally leaks on his Instagram Live that “The song is around four minutes long"

Write a function to classify song length. Possible matches will be between 3.5 and 4.5 minutes. Make a new dataset containing all possibilities from apple music (with the info of time and genre - remember to use your genre from last clue).

```{r}
song_length <- function(minutes) {
  case_when(minutes < 3.5 ~ "Too short",
            minutes <= 4.5 ~ "Possible match",
            minutes > 4.5 ~ "Too long")}

apple_artists <- apple_clean |>
  mutate(duration_match = song_length(track_duration_min)) |>
  filter(genre == "Pop",
         duration_match == "Possible match")

head(apple_artists)
```

### Clue 3: Promo song filter

The hit song played by the mystery artist has medium `energy` (45%-65%) and medium-high `danceability` (60%-75%). Write a data frame function that filters for this, then create a dataset containing all potential artists from Spotify music, using the new function.

```{r}
energy_and_danceability <- function(df) {
  
  df |>
    filter(energy >= 0.45,
           energy <= 0.65,
           danceability >= 0.60,
           danceability <= 0.75)
  
}

spotify_artists <- spotify_clean |>
  energy_and_danceability()

head(spotify_artists)
```

### Clue 4: Combine Datasets

Since such huge names are talking about this mystery artist, you know that the artist has songs on both Apple Music and Spotify. Make a new, combined potential artists list, but remove any duplicates before combining.

::: {.callout-note style="color: green"}
## Hint!

Try using distinct() to remove any duplicates of songs!
:::

```{r}
spotify_artists <- spotify_artists |>
  distinct(track_name, track_artist, .keep_all = TRUE)

apple_artists <- apple_artists |>
  distinct(track_name, artist_name, .keep_all = TRUE)

final_artists <- spotify_artists |>
  inner_join(apple_artists,
             by = join_by(track_name == track_name,
                          track_artist == artist_name,
                          release_year == release_year))
```

### Clue 5: Predict Popularity

TMZ leaks that the song is popular. The mystery song is more popular than expected based on audio features of `danceability`, `energy`, `tempo`, and `acousticness`. Use linear regression to find the popular artists. Make sure that you only keep the top three in popularity.

::: {.callout-note style="color: green"}
## Hint!

You will want to use the predict() function, find more information about predict() here: https://www.geeksforgeeks.org/r-machine-learning/predict-function-in-r/
:::

```{r}
popularity_model <- lm(track_popularity ~ danceability + energy + tempo + acousticness,
                       data = final_artists)

final_artists <- final_artists |>
  mutate(predicted_popularity = predict(popularity_model, newdata = final_artists),
         popularity_difference = track_popularity - predicted_popularity) |>
  slice_max(popularity_difference, n = 3)

head(final_artists)

#if we didn't learn how to mutate this stuff like predict() or something, maybe put a callout or a website link to assist
```

### Clue 6: Going International

A fan notices that all of these hints were posted at some strange times in Pacific Time. You suspect that the mystery artist is actually NOT from the Unites States. You notice that most of these hints are posted when the UK is up and at 'em. Join the remaining potential artists to the global artist data and filter for artists from the United Kingdom.

```{r}
final_artists <- final_artists |>
  left_join(global_clean,
            by = join_by(track_artist == artist_name)) |>
  filter(str_detect(country, "United Kingdom"))

head(final_artists)
```

### Final Artist

You're almost there! Drake tweets "Yoo, this Cal Poly headliner got me mad nostalgic, reminds me of my 'Kiki do you love me' days. The promo song is one of the best songs from the 2010s. Complete the final filter based on these clues, and display your results using the gt() function.

```{r}
final_artists |>
  filter(release_year >= 2010,
         release_year < 2020) |>
  select(track_name,
         track_artist) |>
  gt() |>
  tab_header(title = "Mystery Concert Artist and Big Song") |>
  cols_label(track_name = "Song",
             track_artist = "Artist")
```

::: callout-note
## Result

The artist and their headlineing song is ... \_\_\_\_One Direction headlining with Night Changes\_\_\_\_!!
:::
