---
title: "STAT-331-Final-Project"
date: now
author: "Mia Sahagian, Suhani Kashyap, Hanna Karam"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
editor: source
execute: 
  echo: true
  message: false
  warning: false
---

## Style to dos

```{r}
# Rewrite all questions + make it more fun like this artists said this type pls
# Make sure all questions include a clear description on what to do for that question
# I lowk labeled it suspects from instinct but maybe a different word would be better so maybe rename data
#but if you rename the data make sure it gets changed everywhere
# Add any callouts/that purple thing that looks like code out of code
# Check with her labs to see if anything else should be included for style
#pls check my code and logic
#I just did displays with head() for logic but we can change this idk, kable instead or nothing idk
#when I say high danceability or whatever we need to provide a range instead/also in the question
```

## Packages

```{r}
library(tidyverse)
library(knitr)
library(janitor)
library(gt)
```

# Mystery Concert

**Background**: A mystery music artist is organizing a concert at Cal Poly. The biggest celebrities in the world are dropping hints on social media about who is behind it. As a journalist at Mustang News, you need to follow the hints across Spotify, Apple Music, and global information to discover the mystery artist and let the students at Cal Poly know who’s performing.

## Data Dictionary & Import

#### Global Music:

This dataset includes 100,000+ music artists from around the world, recording names, genres (pop, rock, jazz, or classical), profile image (in a link), and location (where the artist is from).

Find more info on the data here: <https://www.kaggle.com/datasets/harshdprajapati/worldwide-music-artists-dataset-with-image>

#### Apple Music:

The dataset includes information on various artists/tracks such as artist name, track name, genre, track duration, and release date, among others.

Find more info on the data here: <https://www.kaggle.com/datasets/kanchana1990/apple-music-dataset-10000-tracks-uncovered>

#### Spotify Music:

The dataset includes information about songs such as artist name, album name and release date, as well as audio features such as tempo, danceability, acoustiness, energy, and more.

Find more info on the data here: <https://www.kaggle.com/datasets/solomonameh/spotify-music-dataset?resource=download>

```{r}

global_music <- read.csv("./Global Music Artists.csv")

apple_music <- read.csv("./apple_music_dataset.csv")

spotify_music <- read.csv("./high_popularity_spotify_data.csv")
```

## Data Cleaning

### 1. Clean the Apple Music Data

Make the dataset apple_clean and keep only the variables needed for the investigation: artist name, track name, primary genre, release date, and track duration in milliseconds. There should be no columns related to censorship, explicitness, and price. Edit to have release date in year month day hours min second format; make release year be extracted from date, and track time be in minutes.

```{r}
apple_clean <- apple_music |>
 select(artist_name = artistName,
        track_name = trackName,
        genre = primaryGenreName,
        release_date = releaseDate,
        track_time_millis = trackTimeMillis) |>
 mutate(release_date = ymd_hms(release_date),
        release_year = year(release_date),
        track_duration_min = track_time_millis / 60000)
```

### 2. Clean the Spotify Music Data

Make the dataset spotify_clean and keep only variables related to artist identity, popularity, release date, and audio features. Keep only variables related to artist identity, popularity, release date, and audio features. Create a new column called "release year" that has the year the track was released. Dont include track_album_release_date.

::: callout-note
## Hint!

Look into parse_date_time() !
:::

```{r}
spotify_clean <- spotify_music |>
 select(track_name,
        track_artist,
        track_popularity,
        track_album_name,
        track_album_release_date,
        playlist_genre,
        danceability,
        energy,
        tempo,
        acousticness) |>
 mutate(track_album_release_date = parse_date_time(track_album_release_date, 
                                                   orders = c("ymd", "ym", "y")),
        release_year = year(track_album_release_date)) |>
 select(-track_album_release_date)
```

### 3. Clean the Global Music data

Make the dataset global_clean and keep only artist name, country, and genre

```{r}
global_clean <- global_music |>
 select(artist_name,
        country,
        artist_genre)
```

#### Check in: Check sizes

Display with... Apple should be 10000 by 7, spotify 1686 by 10, and global 101375 by 3.

```{r}
kable(tibble(dataset = c("Apple Music", "Spotify Music", "Global Music"),
             rows = c(nrow(apple_clean), nrow(spotify_clean), nrow(global_clean)),
             columns = c(ncol(apple_clean), ncol(spotify_clean), ncol(global_clean))))

```

## Follow the clues

### Clue One

A celebrity posts "The mystery artist is definitely from the most common genre".

Find the most common genres in the Apple Music data by creating a plot and sorting the top 10 genres from least common to most common. Which genre is most common?

```{r}
apple_clean |>
 count(genre, sort = TRUE) |>
 slice_max(n, n = 10) |>
 ggplot(aes(x = fct_reorder(genre, n), y = n, fill = fct_reorder(genre, n))) +
 geom_col() +
 labs(title = "Most Common Apple Music Genres",
      x = "Genre",
      y = "Number of Tracks",
      fill = 'Genre') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 25))

#labels from both x-axis and legend...? include or no? Should this be a gradient instead?
```

The most common music genre in the Apple music data is Pop.

### Clue Two

Justin Bieber leaks: “The song is around four minutes long"

Write a function to classify song length. Possible matches between 3.5 and 4.5 min (or like 1min possible difference from Justin). Make a new dataset containing all potential suspects from apple music (with the info of time and genre - remember to use your genre from last clue).

```{r}
song_length <- function(minutes) {
  case_when(minutes < 3.5 ~ "Too short",
            minutes <= 4.5 ~ "Possible match",
            minutes > 4.5 ~ "Too long")}

apple_suspects <- apple_clean |>
  mutate(duration_match = song_length(track_duration_min)) |>
  filter(genre == "Pop",
         duration_match == "Possible match")

head(apple_suspects)
```

### Clue 3

The hit song played by the mystery artist has medium energy and medium-high danceability. Write a data frame function that filters for this, then make a new dataset containing all suspects from spotify music, using the new function.

```{r}
energy_and_danceability <- function(df) {
  
  df |>
    filter(energy >= 0.45,
           energy <= 0.65,
           danceability >= 0.60,
           danceability <= 0.75)
  
}

spotify_suspects <- spotify_clean |>
  energy_and_danceability()

head(spotify_suspects)
```

### Clue 4

The artist has songs on both Apple Music and Spotify. Make a new, combined suspect list, but remove any duplicates before combining. (maybe callout distinct or something about joining by multiple things idk)

```{r}
spotify_suspects <- spotify_suspects |>
  distinct(track_name, track_artist, .keep_all = TRUE)

apple_suspects <- apple_suspects |>
  distinct(track_name, artist_name, .keep_all = TRUE)

final_suspects <- spotify_suspects |>
  inner_join(apple_suspects,
             by = join_by(track_name == track_name,
                          track_artist == artist_name,
                          release_year == release_year))
```

### Clue 5

TMZ leaks the song is popular. The mystery song is more popular than expected based on audio features of ... Use linear regression to find. Keep top 3 in popularity

```{r}
popularity_model <- lm(track_popularity ~ danceability + energy + tempo + acousticness,
                       data = final_suspects)

final_suspects <- final_suspects |>
  mutate(predicted_popularity = predict(popularity_model, newdata = final_suspects),
         popularity_difference = track_popularity - predicted_popularity) |>
  slice_max(popularity_difference, n = 3)

head(final_suspects)

#if we didn't learn how to mutate this stuff like predict() or something, maybe put a callout or a website link to assist
```

### Clue 6

A fan notices that the hints were posted when the UK is up and at em. Join the remaining suspects to the global artist data and filter for artists from the United Kingdom.

```{r}
final_suspects <- final_suspects |>
  left_join(global_clean,
            by = join_by(track_artist == artist_name)) |>
  filter(str_detect(country, "United Kingdom"))

head(final_suspects)
```

### Final Artist

\_\_ says 'this is one of the best songs from the 2010s / the headlining song is the ya anyways. Display the final artist use gt

```{r}
final_suspects |>
  filter(release_year >= 2010,
         release_year < 2020) |>
  select(track_name,
         track_artist) |>
  distinct() |>
  gt() |>
  tab_header(title = "Mystery Concert Artist and Big Song") |>
  cols_label(track_name = "Song",
             track_artist = "Artist")
```

::: callout-note
## Result

The artist and their headlineing song is ...!!\_\_\_\_
:::
