---
title: "STAT-331-Final-Project"
date: now
author: "Mia Sahagian, Suhani Kashyap, Hanna Karam"
format: 
  html:
    embed-resources: true
    code-tools: true
    toc: true
editor: source
execute: 
  echo: true
  message: false
  warning: false
---

# PULL commit PUSH

# Mystery Concert

**Background**: A mystery music artist is organizing a concert at Cal Poly. The biggest celebrities in the world are dropping hints on social media about who is behind it. As a music fanatic journalist at Mustang News, you need to follow the hints across Spotify, Apple Music, and global information to discover the mystery artist and let the students at Cal Poly know who’s performing.

## Data Dictionary & Import

#### Global Music:

This dataset includes 100,000+ music artists from around the world, recording names, genres (pop, rock, jazz, or classical), profile image (in a link), and location (where the artist is from).

Find more info on the data here: <https://www.kaggle.com/datasets/harshdprajapati/worldwide-music-artists-dataset-with-image>

#### Apple Music:

The dataset includes information on various artists/tracks such as artist name, track name, genre, track duration, and release date, among others.

Find more info on the data here: <https://www.kaggle.com/datasets/kanchana1990/apple-music-dataset-10000-tracks-uncovered>

#### Spotify Music:

The dataset includes information about songs such as artist name, album name and release date, as well as audio features such as tempo, danceability, acoustiness, energy, and more.

Find more info on the data here: <https://www.kaggle.com/datasets/solomonameh/spotify-music-dataset?resource=download>

```{r}

global_music <- read.csv("./Global Music Artists.csv")

apple_music <- read.csv("./apple_music_dataset.csv")

spotify_music <- read.csv("./high_popularity_spotify_data.csv")
```

## Packages

```{r}
#| label: packages

library(tidyverse)
#library(knitr) add a gt kableextra type beat
```

## Data Cleaning

```{r}
apple: artist name, releasedate, trackname, primarygenre, tracktimemillis
rename things in apple (remove: anything with Censored, explicitness, any price)

global:artist_name and country

spotify: track_artist, energy, tempo, danceibility, track album release date, track popularity, acoustiness
(playlist genre with primary genre?)(remove: anything with id, anything with playlist)
```

### Clean the Apple Music Data

Keep only the variables needed for the investigation: artist name, track name, primary genre, release date, and track duration in milliseconds. There should be no columns related to censorship, explicitness, and price.

```{r}
apple_clean <- apple_music |>
 clean_names() |>
 select(
   artist_name = artistName,
   track_name = trackName,
   primary_genre_name = primaryGenreName,
   release_date = releaseDate,
   track_time_millis = trackTimeMillis) |>
 mutate(
   release_date = ymd_hms(release_date),
   release_year = year(release_date),
   track_duration_min = track_time_millis / 60000
 )
```

### Clean the Spotify Music Data

Keep only variables related to artist identity, popularity, release date, and audio features.

```{r}

```

### Ideas

Pick: Night Changes - OneDirection

Lubridate time zone to figure out they're from the UK (ex it's this time for the artist)

Fix the track_name/merge/get ride of playlist info

Don't join apply music country

1\) Data cleaning stuff (which rows to save/not save)(how many rows and cols)

2\) Clues

3\) GGPlot most popular genre, factor reorder by genre (Pop)

4\) Energy level: medium high danceability (.672) and medium energy (0.52) leaked by justin bieber

6\) song duration - tracktimemillis

5\) data joins

6\) Country of origin/when track came out/release date - with date
